---
title: Day One Review
---

```{r include=FALSE}
source('common.R')
```

##

| Start        | End      | Duration | What                                                | Who      |
|-------------:|---------:|:--------:|:----------------------------------------------------|:--------:|
|  1:00 PM     |  2:00 PM	|  0:60	   | [Review](02_01_day_one_review.html)                 | Brian    | 
|  2:00 PM     |  3:00 PM	|  0:60    | [`dplyr`](02_02_TBD.html)                           | Adam     | 
|  3:00 PM     |  3:30 PM	|  0:30    | Break	                                             | You      | 
|  3:30 PM     |  4:15 PM	|  0:45	   | [More visualization](02_03_more_visualization.html) | Brian    | 
|  4:15 PM     |  5:00 PM	|  0:45	   | Tree-based methods                                  | Adam     | 

# Before we begin

## Did they bother to prepare?

Student: Psst, Adam/Brian. You didn't cover X, Y or Z.  

Adam/Brian: We didn't cover O, P, or Q either. R is VAST.

## Be easy on yourself

* Mistakes are very good things. 
  * If you don't make mistakes, you may not be trying enough new stuff.
  * Brian has made ALL of the mistakes
* There are often multiple ways to do something in R.
  * This is a good thing, except when it isn't.
* Don't reinvent the wheel. Your use case is probably not rare. 
  * Need the first day of a quarter? There's a function for that.
  * Need way to translate data into sound? There's a package for that.
  * Need to run a chain ladder, Dave Clark growth curve, hierarchical linear model for a loss reserve? There's a package for that.

## Points of emphasis

* Everything is vectorised. This is essential - but not hard! - to understand.
* Use projects. They're low cost/high value.
* R vs. Excel/SAS/Whatever
  * Excel: Data is visible, logic is hidden.
  * R: Data is hidden, logic is visible.
* Horses for courses
  * SAS is faster, but expensive.
  * Excel can format printed output, but can't do cluster analysis.
  * SQL can extract data, but can't draw pictures.
  * Be a Swiss Army actuary!

# Stuff we might or might not have gone over

## Comments and strings

* Comments are made with a hash mark: `#`
* Comments should explain _why_, not _what_
  * If "what" isn't obvious, your code might need to be rewritten
* CTRL+SHIFT+C in RStudio

## Strings

Single quote, double-quote. Whatever. Just use the same for the start and end of the string.

```{r eval = FALSE}
"This is the name of someone"
'This is also the name of someone'
"This is someone's name"
'This is also someone''s name'
```

## Useful functions

* `sort()`
* `paste0()`
* `order()`
* `which()`
* `table()`

## `sort()`

```{r}
people <- c('Brian', 'Adam', 'Kellee', 'Leanne')
characteristics <- c('cranky', 'jolly', 'jolly', 'jolly')
sort(people)
sort(characteristics)
```

## `order()` vs. `sort()`

* `order()` returns indices
* `sort()` returns items

```{r}
order(people)
people[order(people)]
sort(people)
```

## `which()`

```{r}
which(people == 'Adam')

happy_people <- which(characteristics != 'cranky')
people[happy_people]
```

## `table()`

```{r}
table(characteristics)
```

# Programmatic stuff

## Creating functions

```{r}
add_two <- function(thing_one, thing_two){
  thing_one + thing_two
}

add_two(3, 4)
```

## Functions: loads more nuance

* Default arguments
* Argument checking
* Early return
* `...`

## Loops

```{r}
for (i in seq_len(5)) {
  print(i)
}

for (i_name in seq_along(people)) {
  print(people[i_name])
}
```

## Loops: don't use them

* `lapply()`, and `map()` do the same thing
* More expressive, less code
* Easier to pipe
* Parallel processing

##

```{r}
library(tidyverse)

count_e <- function(x) {
  str_count(x, '[Ee]')
}
```

##

`map(my_vector, my_function)`

```{r}
lapply(people, count_e)

map(people, count_e)
map_int(people, count_e)
```

## Loops: you rarely need them

Vectorized operations _are_ loops

```{r}
for (i_name in seq_along(people)) {
  people[i] <- tolower(people[i])
}

people <- toupper(people)
```

## When loops are appropriate

* Conditional execution - do until
* Recursive calculation - can't run in parallel
* You're tired and you just want to move on with your life - beware of technical debt

## Conditional execution

```{r eval = FALSE}
if (my_value == something) {
  print("My value equals something")
}

if (my_value == something) {
  print("My value equals something")
} else if (my_value == something_else) {
  print("My value equals something else")
} else ()
```

## `if` vs. `ifelse()`

```{r}
if (people == "Adam") {
  print("Hi Adam!")
}

ifelse(
  people == "Adam"
  , 'In Connecticut'
  , 'Somewhere else'
)
```


# The exercises

## Demo

1. Create a project to use during the bootcamp.
1. Create a folder in that project called 'data'.
3. Create a script which executes the line `1 + 1`.
4. Execute that script.

# Data import/export

1. Create a list of names, and birth dates in Excel. Read that data into R.
2. Add one year to the dates, and save the results into a .CSV file.
3. Read in both tables and verify that your data arithmetic is sound.